---
title: "COVID-19 Prediction Flexdashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: embed
runtime: shiny
---

```{r global, include=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
library(flexdashboard); library(plotly); library(shiny);
library(e1071); library(ggplot2); 
library(tidyverse); 
library(rvest);
library(highcharter); 
# load data
dat <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")


# delete redundant columns
datnew <- dat[-c(1:6,8:11)] 

# add case number 
statecase <- as.data.frame(datnew %>%                                      
  group_by(Province_State) %>%                      
  summarise_if(is.numeric,funs(sum)))
# delete princess *2 
statecase <- statecase[-c(3,10,14),]  
date <- names(statecase)[-c(1)]
lod <- length(date)  # length of date 
time_order_now <- c(1:lod) # 1,2,3... the last day of record
dif <- 7 # days in future
future_forecast <- c(1:(lod+dif)) # 1,2,3...last day of prediction
adjusted_dates <- c((lod+1):(lod+dif)) # prediction date 
# state name
state_name <- as.character(statecase$Province_State)

# us map
cov19_us <- read_html('https://www.worldometers.info/coronavirus/country/us/')

df <- cov19_us %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table()  %>% as.data.frame()

df$TotalCases <- gsub(",", "",df$TotalCases)
#df <- df[1:51,]
df <- df[-c(nrow(df)),]

df_plot <- df[2:56,1:2]
df_plot["lognumber"] <- log(as.numeric(df_plot$TotalCases))
df_plot["knumber"] <- round(as.numeric(df_plot$TotalCases)/1000,digits=1)
mapdata <- get_data_from_map(download_map_data("countries/us/us-all"))
glimpse(mapdata)

# world 
cov19_world <- read_html('https://www.worldometers.info/coronavirus/')

df_world <- cov19_world %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table()  %>% as.data.frame()

df_world <- df_world[nrow(df_world),]
```



United States
=======================================================================

Column {.sidebar data-width=324}
-----------------------------------------------------------------------
Last Updated on:
```{r}
rmarkdown::render_delayed({div(Sys.time())})
```

```{r}
selectInput('state', 'State', state_name,selected = NULL)
```

```{r}
hcmap("countries/us/us-all", data = df_plot, value = "knumber",
      joinBy = c("name", "USAState"), name = "Total Cases",
      dataLabels = list(enabled = TRUE, format = '{point.name}'),
      borderColor = "#000000", borderWidth = 1,
      tooltip = list(valueSuffix = " k"))
       
```


Column {.tabset}
-----------------------------------------------------------------------
### Prediction
```{r}
svmmodel <- function(input_state_name,state_name,time_order_now,
                    statecase,future_forecast,lod,dif){
  index <- match(input_state_name,state_name)
  x <- time_order_now
  y <- as.numeric(statecase[index,1:lod+1])
  model <- svm(x, y,kernel="polynomial") 
  pred <- predict(model, future_forecast)
  # parameter 
  error <- y - pred[1:lod]
  svm_error <- sqrt(mean(error^2))
  svm_tune <- tune.svm(time_order_now,y,kernel = "polynomial",
                       gamma=c(0.1, 1, 5),
                       coef0 = c(0.05,0.1, 1, 5))
  best.svm = svm_tune$best.model
  newprediction = predict(best.svm, future_forecast)
  error_new <- y - newprediction[1:lod]
  svm_error_new <- sqrt(mean(error_new^2))
  
  # create dataframe to plot
  date_plot_true <- seq(as.Date("2020/1/22"), by = "day", length.out = lod)
  date_plot_pred <- seq(as.Date("2020/1/22"), by = "day", length.out = lod+dif)
  
  truevalue <- data.frame(
    date_plot_true,
    y)
  
  prediction <- data.frame(
    date_plot_pred,
    newprediction)
  
  cols1 = c("dates", "Confirmed number")
  cols2 = c("dates","SVM Prediction")
  colnames(truevalue) = cols1
  colnames(prediction) = cols2
  
  # glm model

  #find first positive case
  stateweneed = input$state
  
  state_confirmed = filter(datnew, Province_State == stateweneed)[-1] %>% colSums() %>% as.data.frame()
  names(state_confirmed) = c('confirmed')
  
  #find first positive case
  first_positive = which(state_confirmed$confirmed > 0)[1]
  
  time = as.numeric(seq(first_positive,nrow(state_confirmed)))
  
  dat = data.frame(time, (time - first_positive + 1),
                   state_confirmed[first_positive:nrow(state_confirmed),])
  names(dat) = c('days since 1.22.20', 'days_since_first_case','cumulative_confirmed')
  
  #train model 
  glm1 = glm(cumulative_confirmed ~ log(days_since_first_case), 
             data = dat,
             family= quasipoisson(link = 'log'))
  
  #make prefiction
  predicted_time = as.numeric(seq(first_positive,nrow(state_confirmed)+7))
  predicted_case = exp(glm1$coefficients[1]+log(predicted_time-first_positive+1)*glm1$coefficients[2])
  
  # create dataframe of glm to plot
  first_positive_date <- as.character(seq(as.Date("2020/1/22"), 
                                          by = "day", length.out = first_positive)[first_positive])
  glm_date_seq <- seq(as.Date(first_positive_date), by = "day", length.out = lod+dif-first_positive+1)
  glmprediction <- data.frame(
    glm_date_seq,
    round(predicted_case,digits=0)
  )
  
  col3 = c("dates","GLM Prediction")
  colnames(glmprediction) <- col3
  
  # create combine dataframe
  combinedf <- right_join(truevalue,prediction,by="dates")
  combinedf <- right_join(glmprediction,combinedf,by="dates")
  combinedfmelted <- reshape2::melt(combinedf,id.var="dates")
  
  return(combinedfmelted)
}

renderPlotly({
 
  g <- ggplot(data=svmmodel(input$state,state_name,time_order_now,
                              statecase,future_forecast,lod,dif),
            aes(x=dates,y=value,col=variable)) + 
  geom_line(size=1) + 
  labs(x = "Date",
       y = "Cumulative Confirmed Cases",
       color = "Legend") +
  ggtitle(paste("Projected Trend of Confirmed Covid-19 Case in",input$state)) +
  theme(
    plot.title = element_text(size=12, face="bold",hjust=0.5),
    axis.title.x = element_text(size=8, face="bold"),
    axis.title.y = element_text(size=8, face="bold")
  )
  
  ggplotly(g,dynamicTicks=TRUE)
})

```

### US data
```{r}
renderTable({
  lastday <- statecase[,c(1,(lod-5):(lod+1))]
  col2 <- as.character(seq(as.Date("2020/1/22"), by = "day", length.out = lod)[(lod-6):(lod)])
  colnames(lastday) <- c("State",col2)
  lastday
})
```

World Data
=======================================================================

Row
-------------------------------------

### Summary of coronavirus data aroung the globe
```{r}
renderTable({
  
cov19_world <- read_html('https://www.worldometers.info/coronavirus/')

df_world <- cov19_world %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table()  %>% as.data.frame()
})
```


About
=======================================================================

Column {.tabset}
-------------------------------------
### Flexdashboard Overview

**Data Acquisition**

The raw data used for model training is pulled from [here](https://github.com/CSSEGISandData/COVID-19) COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University every time the dashboard refreshes.

**Model Development**

The flexdashboard contains two prediction models forecasting the confirmed Covid-19 cases for the week ahead. 

* Support Vector Machine (SVM) Regression Model

Apply a SVR model with a polynomial kernel to fit the cumulative confirmed case data. Tune model parameters using trial and error. Use the optiaml SVR model to predict the confirmed case for the week ahead.



* GlM Model

A quasipoisson model was applied to predict the trend of confirmed Covid-19 cases for the week ahead. It is assumed tha the response variable Y (cumulative confirmed cases) follows a poisson distribution, and the logrithm of its expected value can be modeled by a linear combinatino of parameters (the log value of days since first positive Covid-19 case confirmed in that state).


$$log(E[Y_i | X_i = x_i]) = log(\mu_i) = \beta_0 + \beta_1x_i$$


**Interactive Component and Visualization**

Select which state of interest from the sidebar on the right to see the projected confirmed Covid cases for the week ahead, and the dashboard will display the cumulative cases of Covid-19 of that state since Jan,22,2020 and two predicted trends for the week ahead by the two models mentioned above.


Hover over the hcmap at the left bottom corner to see the confirmed cases of 50 states. Values are in k scale. Darker the color, more confirmed cases of Covid-19.


### Disclaimer

This flexdashboard was made as a final project for the course the JHU Data Science for Biomedical Engineering taught by Brian Caffo. This is not official health information. Under no circumstances should this app be used for medical guidance. 


Please refer to official CDC and WHO website for recommendations and guidance.

CDC

https://www.cdc.gov/coronavirus/2019-ncov/index.html

WHO

https://www.who.int/emergencies/diseases/novel-coronavirus-2019


### References
Data (prediction model)

https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv


Data (domestic/USA)

https://www.worldometers.info/coronavirus/country/us/

Data (worldwide)

https://www.worldometers.info/coronavirus/

CDC travel restrictions

https://www.cdc.gov/coronavirus/2019-ncov/travelers/map-and-travel-notices.html











